Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-10-22 forecast/paper/paper.tex
Sync up before branch
Make suggested minor changes to fcast

\begin{abstract}
  Reliable, short-term forecasts of traditional public health reporting streams
  (such as cases, hospitalizations, and deaths) are a key ingredient in
  effective public health decision-making during a pandemic. Since April 2020,
  our research group has worked with data partners to collect, curate, and make
  publicly available numerous real-time COVID-19 indicators, providing multiple
  views of pandemic activity.  This paper studies the utility of these
  indicators from a forecasting perspective. We focus on five indicators,
  derived from medical insurance claims data, web search queries, and online
  survey responses. For each indicator, we ask whether its inclusion in a simple
  model leads to improved predictive accuracy relative to a similar model
  excluding it. We consider both probabilistic forecasting of confirmed COVID-19 
  case rates and binary prediction of case ``hotspots''.  Since the values of 
  indicators (and case rates) are commonly revised over time, we take special
  care to ensure that the data provided to a forecaster is the version that
  would have been available at the time the forecast was made.   
  % We demonstrate how less careful backtesting can lead to misleading
  % retrospective evaluations.   
  Our  analysis shows that consistent but modest gains in predictive accuracy 
  are obtained by using these indicators, and furthermore, these gains are
  related to periods in which the auxiliary indicators behave as ``leading
  indicators'' of case rates.     
\end{abstract}

@@

\dropcap{T}racking and forecasting indicators from public health reporting
streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
crucial for understanding disease spread, formulating public policy responses,
and anticipating future public health resource needs.  A companion paper
\cite{Reinhart:2021} describes our research group's (Delphi's) efforts in
curating and maintaining a database of real-time indicators that track COVID-19
activity and other relevant phenomena. The signals (a term we use synonomously
with ``indicators'') in this database are accessible through the COVIDcast API
\cite{CovidcastAPI}, with associated R \cite{CovidcastR} and Python
\cite{CovidcastPy} packages for convenient data fetching and processing
tools. In the current paper, we aim to quantify the utility provided by a core
set of these indicators for two fundamental prediction tasks: probabilistic
forecasting of COVID-19 case rates and prediction of future COVID-19 case
hotspots (defined by the event that a relative increase in COVID-19 cases
exceeds a certain threshold).

@@

Lastly, our treatment of the Google-AA signal is different from the rest.
Because Google's team did not start publishing this signal until early
September, 2020, we do not have true vintage data before then.  Furthermore, the
latency of the signal was always at least one week through 2020.
% However, this signal is never revised after initial publication (confirmed via
% personal communication with the Google team that produces this signal)
However, unlike the claims-based signals, there is no reason for revisions to
occur after initial publication, 
and furthermore,
the latency of the signal is not an unavoidable property of the data type, so we
simply use finalized signal values, with zero latency, in our analysis.

@@

In our analysis, we let the prediction date $t$ run over each day in between
early/mid June and December 31, 2020.  The precise start date differs for 
forecasting and hotspots prediction; for each task it was chosen to be the
earliest date at which the data needed to train all models was available, which
ends up being (per our setup, with 21 days of training data and lagged values 
of signals for features, as we will detail shortly) June 9, 2020 for
forecasting, and June 16, 2020 for hotspot prediction. (The bottleneck here is
the CTIS-CLI-in-community signal, which does not exist before early April 2020,
when the survey was first launched). The end date was chosen again with a 
consideration to align both tasks as best as possible, and because very few
hotspots exist for the several few months of 2021, due to the general and
gradual decline of the pandemic during this period.

@@

\item In this same analysis, when an indicator model performs better than AR in
  a decreasing period, this tends to co-occur with instances in which the
  indicator ``leads'' case rates (meaning, roughly, on a short-time scale in a
  given location, its behavior mimics that of future case rates).  On
  the other hand, if an an indicator model does better in periods of
  increase, or worse in periods of increase or decrease, its performance is
  not as related to leadingness.   
\end{itemize}

@@

To test this hypothesis, we reran all of the prediction experiments but with
$X_{\ell,t}$ in the each indicator models replaced with suitable random noise
(bootstrap samples from a signal's history).  The results, shown and explained
more precisely in the Supplementary Materials, are vastly different (worse) than
the original set of results.  In both forecasting and hotspot prediction, the
``fake'' indicator models offered essentially no improvement over the pure AR
model, which strongly rejects (informally speaking) the implicit regularization
hypothesis.

>>>>

\begin{abstract}
  Reliable, short-term forecasts of traditional public health reporting streams
  (such as cases, hospitalizations, and deaths)
  ought to be a key
  % are a key
  ingredient in
  effective public health decision-making during a pandemic. Since April 2020,
  our research group has worked with data partners to collect, curate, and make
  publicly available numerous real-time COVID-19 indicators, providing multiple
  views of pandemic activity.  This paper studies the utility of these
  indicators from a forecasting perspective. We focus on five indicators,
  derived from medical insurance claims data, web search queries, and online
  survey responses. For each indicator, we ask whether its inclusion in a simple
  model leads to improved predictive accuracy relative to a similar model
  excluding it. We consider both probabilistic forecasting of confirmed COVID-19 
  case rates and binary prediction of case ``hotspots''.  Since the values of 
  indicators (and case rates) are commonly revised over time, we take special
  care to ensure that the data provided to a forecaster is the version that
  would have been available at the time the forecast was made.   
  % We demonstrate how less careful backtesting can lead to misleading
  % retrospective evaluations.   
  Our  analysis shows that consistent but modest gains in predictive accuracy 
  are obtained by using these indicators, and furthermore, these gains are
  related to periods in which the auxiliary indicators behave as ``leading
  indicators'' of case rates.     
\end{abstract}

@@

\dropcap{T}racking and forecasting indicators from public health reporting
streams---such as confirmed cases and deaths in the COVID-19 pandemic---is
crucial for understanding disease spread, correctly formulating public policy responses,
and rationally planning future public health resource needs.  A companion paper
\cite{Reinhart:2021} describes our research group's (Delphi's) efforts in
curating and maintaining a database of real-time indicators that track COVID-19
activity and other relevant phenomena. The signals (a term we use synonomously
with ``indicators'') in this database are accessible through the COVIDcast API
\cite{CovidcastAPI}, with associated R \cite{CovidcastR} and Python
\cite{CovidcastPy} packages for convenient data fetching and processing
tools. In the current paper, we aim to quantify the utility provided by a core
set of these indicators for two fundamental prediction tasks: probabilistic
forecasting of COVID-19 case rates and prediction of future COVID-19 case
hotspots (defined by the event that a relative increase in COVID-19 cases
exceeds a certain threshold).

@@

Lastly, our treatment of the Google-AA signal is different from the rest.
Because Google's team did not start publishing this signal until early
September, 2020, we do not have true vintage data before then.  Furthermore, the
latency of the signal was always at least one week through 2020.  However,
unlike the claims-based signals, there is no reason for revisions to occur after
initial publication, and furthermore, the latency of the signal is not an
unavoidable property of the data type, so we simply use finalized signal values,
with zero latency, in our analysis.

@@

In our analysis, we let the prediction date $t$ run over each day in between
early/mid June and December 31, 2020.  The precise start date differs for 
forecasting and hotspots prediction; for each task it was chosen to be the
earliest date at which the data needed to train all models was available, which
ends up being (per our setup, with 21 days of training data and lagged values 
of signals for features, as we will detail shortly) June 9, 2020 for
forecasting, and June 16, 2020 for hotspot prediction. (The bottleneck here is
the CTIS-CLI-in-community signal, which does not exist before early April 2020,
when the survey was first launched). The end date was chosen again with a 
consideration to align both tasks as best as possible and to coincide with the
introduction of widely available vaccines.

@@

\item In this same analysis, when an indicator model performs better than AR in
  a decreasing period, this tends to co-occur with instances in which the
  indicator ``leads'' case rates (meaning, roughly, on a short-time scale in a
  given location, its behavior mimics that of future case rates).  On
  the other hand, if an indicator model does better in periods of
  increase, or worse in periods of increase or decrease, its performance is
  not as related to leadingness.   
\end{itemize}

@@

To test this hypothesis, we reran all of the prediction experiments but with
$X_{\ell,t}$ in each indicator model replaced with suitable random noise
(bootstrap samples from a signal's history).  The results, shown and explained
more precisely in the Supplementary Materials, are vastly different (worse) than
the original set of results.  In both forecasting and hotspot prediction, the
``fake'' indicator models offered essentially no improvement over the pure AR
model, which---informally speaking---strongly rejects the implicit regularization
hypothesis.
